DocSplit: Simple Contrastive Pretraining for \\ Large Document Embeddings

### 2023-06-20 emnlp2023-latex/paper.tex
update

% This is also not strictly necessary and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
%\usepackage{inconsolata}

\usepackage{microtype}
\usepackage{enumitem}
\usepackage{times}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{color}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb,amsthm}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{caption}
%\usepackage{paralist}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
\newcommand{\todo}[1]{{\color{red}{#1}}}
\newcommand{\our}{\mbox{\textsc{SimDE}}}
% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{SimDE: Simple Contrastive Learning for Document Embeddings}

@@

\begin{document}
\maketitle
\begin{abstract}
%High-quality text embeddings are essential to various natural language understanding tasks especially for low-resource settings. 
%Existing methods are proposed for general text understanding but representations for long text are less explored.
%Basically, to have better embeddings for long text, a model should have the ability to recognize key information in a long text. 
%We introduce the $\our$ 
%To this end, we present \our, a simple unsupervised contrastive learning framework for long text embeddings in this paper. 
%Specifically, we pretrain a language model to distinguish if two texts have the same topic without any supervision. 
%The positive pairs are constructed by our key information redundancy assumption for long text.
% Experimental results on five datasets show that \our~outperforms the state-of-the-art models by $3.9\%$ on classification tasks and $12.0\%$ macro-F1 in average respectively under general and few-shot text classification settings.

% This paper presents $\our$, the first unsupervised pretraining
% method designed specifically for documents.
% $\our$ uses the contrastive learning framework, and our main contribution is a simple but effective data augmentation technique for generating similar text pairs.
%A key advantage of $\our$ over other techniques for working with
%long text is that it requires no changes to model architecture, and so it is widely applicable.
% Specifically, we pretrain a language model to distinguish if two texts have the same topic without any supervision or specific model architectures, and so it is widely applicable.
%We use $\our$ pretraining to improve the embeddings of previously
%developed state-of-the-art models.
% The positive pairs are constructed by our key information redundancy assumption for documents.
% On standard classification datasets, $\our$ improves all baseline
% models, with an average improvement of $3.9\%$ macro F1 score.
% We also consider a few-shot setting where we show an average
% improvement of $12.0\%$.

Existing pretrained transformer models generate high-quality sentence and paragraph embeddings, but do not work well for large documents.
A growing body of research tackles this problem by proposing new model architectures that are more efficient with large context windows, but all of this research still uses the classic token masking strategy for training.
Token masking is a fundamentally \emph{local} operation in that context of the entire document is rarely needed to correctly predict the masked token---the words much closer to the token will have a much larger
influence.
We propose a novel \emph{global} method for training that requires a model to equally consider all portions of a document.
Our Simple Document Embedding (SimDE) method is based on contrastive learning.
Like token masking, SimDE is fully unsupervised and can work with any existing model architecture.
We experimentally evaluate the SimDE on a wide range of architectures on standard classification datasets for large documents.
We find an average improvement of $3.9\%$ macro F1 score in a standard
regime with large training datasets and a massive $12.0\%$ improvement in the macro F1 score in a few-shot setting.
\end{abstract}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{fig/simde.pdf}
\caption{Overall framework of \our. The document is randomly divided into two exclusive subsets of sentences and the two subsets work as positive pairs for contrastive learning. Other instances in the same batch are used as negatives.
}
\label{overall}
\end{figure}

@@

To evaluate the quality of document embeddings, we conduct standard and few-shot text classification on five long text datasets involved in News and scientific articles. 
The experimental results show that \our~with two kinds of model structures (i.e.,~BERT and Longformer) can both achieve significant improvements compared to state-of-the-art baselines.

@@

\subsection{SimDE}
The critical problem in contrastive learning is how to construct positive pairs $(x, x^{+})$.
In representation learning for visual tasks~\cite{Chen2020ASF}, an effective solution is to take two random transformations of the same image (e.g.,~flipping, rotation).
Similarly, in language representations, previous works~\cite{Gao2021SimCSESC, Karpukhin2020DensePR, Meng2021COCOLMCA, Li2022UCTopicUC} apply augmentation techniques such as dropout, word deletion, reordering, and masking.

@@

\begin{table*}
    \centering
% \scalebox{0.7}{
% \setlength{\tabcolsep}{1.mm}{
    \input{fig/result.tex}
    % }}
    \caption{For all performance measures, larger numbers are better. Our pre-trained model achieves the best results in all cases.}
    \vspace{-3mm}
    \label{results}
\end{table*}

@@

\subsection{Datasets}
We use the following classic documents datasets to evaluate our method:
(1) Fake News Corpus~\footnote{https://github.com/several27/FakeNewsCorpus};
(2) 20NewsGroups~\cite{Lang1995NewsWeederLT};
(3) arXiv articles dataset~\footnote{https://www.kaggle.com/datasets/Cornell-University/arxiv};
(4) New York Times Annotated Corpus (NYT) ~\cite{sandhaus2008new}; and
(5) BBCNews~\footnote{http://mlg.ucd.ie/datasets/bbc.html}.
We do not use semantic textual similarity (STS) tasks~\cite{Agirre2012SemEval2012T6} because the sentences in these tasks are short which is not suitable to evaluate long text embeddings.

\subsection{Baselines}
We compare our pre-trained model to the baselines of two groups.
(1) BERT based models include BERT~\cite{Devlin2019BERTPO}, SimCSE~\cite{Gao2021SimCSESC}, CT-BERT~\cite{Carlsson2021SemanticRW}. For a fair comparison, we also train a SimCSE with our pretraining dataset (SimCSE$_\mathrm{long}$).
(2) Transformers specified for long sequences include Longformer~\cite{Beltagy2020LongformerTL} and BigBird~\cite{Zaheer2020BigBT}.
We train two versions of \our~with BERT and Longformer (i.e.,~\our$_\mathrm{bert}$ and \our$_{\mathrm{long}}$) for comparison.
We do not include RoBERTa~\cite{Liu2019RoBERTaAR} and IS-BERT~\cite{Zhang2020AnUS} as our baselines because SimCSE achieves better results than these methods according to the paper.

@@

\textbf{Results.}
Table~\ref{results} shows the evaluation results on different datasets. 
Overall, we can see that \our~achieves the best performance over the 5 document datasets and consistently improves the document embeddings with BERT and Longformer structures. 
Specifically, methods pretrained with contrastive objectives (i.e.,~CT-BERT, SimCSE) outperform general language representations (i.e., BERT) which indicates contrastive objectives designed for text embeddings can largely improve the ability of language models to produce high-quality text embeddings. SimCSE pretrained with our document data (i.e.,~SimCSE$_{\mathrm{long}}$) has similar results as the original SimCSE which indicates simplely increasing the length of pretraining text cannot improve document embeddings.
Compared to SimCSE and Longformer, our model achieves $3.9\%$ and $9.4\%$ average macro-F1 improvements with BERT and Longformer structures respectively. Hence, our contrastive learning method is effective for document embeddings.

@@

\textbf{Results.} 
Table~\ref{results} shows the results of few-shot text classification on these five datasets. 
We can see that \our~(i.e.,~\our$_{\mathrm{bert}}$ and \our$_{\mathrm{long}}$) achieves $12.0\%$ and $24.3\%$ macro-F1 improvements compared to SimCSE and Longformer respectively. These improvements are higher than standard text classification. 
Besides, we also compare the performance of different baselines and \our$_{\mathrm{bert}}$ with different numbers of training instances on 20News. 
The results in Figure~\ref{few_shot} show the improvements from our method become larger as the number of training instances decreases indicating the importance of high-quality document embeddings for low-resource settings. 
Furthermore, our method achieves the best results under different numbers of training instances.

\subsection{Document Retrieval}
We use document retrieval to evaluate the ability of learning the similarity score between two vectors \cite{ Guo2016ADR}. We follow the document retrieval experiment \cite{Tay2020LongRA} and use the ACL Anthology Network (AAN) \cite{Radev2009TheAA} dataset, which identifies if two papers have a citation link, a common setup used in long-form document matching \cite{Jiang2019SemanticTM,Yang2020Beyond5T}.

\textbf{Results.} 
\begin{table*}
    \centering
    \input{fig/document_rerival.tex}
    \caption{}
    \vspace{-3mm}
    \label{retrieval}
\end{table*}
Table~\ref{retrieval} shows the results of document retrieval experiment on AAN dataset.

\section{Conclusion}
\label{sec:conclusion}
In this work, we propose an unsupervised contrastive learning framework for long text embeddings. 
Our method provides a new method for long text data augmentation without any supervision and language models can get large-scale pretraining on any long text. 
We conduct extensive experiments on text classification tasks under fully supervised and few-shot settings.
Results show that our pre-trained model greatly outperforms state-of-the-art text embeddings, especially when the training data is limited.

@@

\textbf{Text Embeddings}. Learning text embeddings is a fundamental task in natural language processing (NLP) which can be used in information retrieval~\cite{Thakur2021BEIRAH}, text similarity~\cite{Gao2021SimCSESC}, classification~\cite{Minaee2021DeepLB}, etc. Existing works focus on learning general text embeddings for different applications. SimCSE~\cite{Gao2021SimCSESC} augment training data by Dropout~\cite{Srivastava2014DropoutAS} to construct positive data for contrastive learning. Sentence-BERT~\cite{Reimers2019SentenceBERTSE} employs siamese and triplet network structures to compare the labeled sentences in the natural language inference (NLI) dataset. Contriever~\cite{Izacard2021UnsupervisedDI} learns text embeddings for information retrieval via contrastive learning considering multilingual corpus. Instructor~\cite{Su2022OneEA} collects a massive dataset from 330 diverse NLP tasks and formulates all tasks into information retrieval tasks with prompts. 
Previous works either rely on massive manually annotated data~\cite{Su2022OneEA, Reimers2019SentenceBERTSE} or focus only on short text (i.e.,~sentences) embeddings~\cite{Gao2021SimCSESC}.
In this paper, we study unsupervised representation learning of document embeddings which usually have longer text than general sentence embeddings and this task brings new challenges for models to understand a document.

@@

\subsection{Attention Weights}
\label{app:attention}
To explore the difference between \our~and other models, we analyze the attention weights of Transformers in different models on the NYT dataset (details in Appendix~\ref{app:attention}). The average weights of different kinds of words are shown in Figure~\ref{attentions}.  We can see that our model has more than $40\%$ higher attention weights on nouns compared to BERT and SimCSE.~\citet{martin-johnson-2015-efficient} shows nouns are more informative than other words in the document understanding. Hence, our pretraining method increases the attention weights of models on nouns which results in higher performance on long text classification.

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{fig/attentions.pdf}
\caption{Attention weights from different models on the NYT dataset.}  
\label{attentions}
\end{figure}

We compute the attention weights for Transformers as follows:
(1) we first extract the attention weights between \texttt{[CLS]} token and all the other tokens;
(2) we compute the averaged weights along different heads in multi-head attention;
(3) the attention weights of the last layer in Transformers are used as the weights for words.
Averaged values are computed for nouns, verbs, adjectives, and other words.

>>>>

% This is also not strictly necessary and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
% \usepackage{inconsolata}

\usepackage{microtype}
\usepackage{enumitem}
\usepackage{times}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{color}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb,amsthm}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{caption}
%\usepackage{paralist}
\usepackage{cleveref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
\newcommand{\todo}[1]{{\color{red}{#1}}}
\newcommand{\our}{\mbox{\textsc{DocSplit}}}
% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{DocSplit: Simple Contrastive Pretraining for \\ Large Document Embeddings}

@@

\begin{document}
\maketitle
\begin{abstract}

Existing model pretraining methods only consider local information.
For example, in the popular token masking strategy, the words closer to the masked token are more important for prediction than words far away.
This results in pretrained models that generate high-quality sentence embeddings, but low-quality embeddings for large documents.
We propose a new pretraining method called $\our$ which forces models to consider the entire global context of a large document.
Our method uses a contrastive loss where the positive examples are randomly sampled sections of the input document, and negative examples are randomly sampled sections of unrelated documents.
Like previous pretraining methods, $\our$ is fully unsupervised, easy to implement, and can be used to pretrain any model
architecture.
Our experiments show that $\our$ outperforms other pretraining methods for document classification, few shot learning, and information retrieval tasks.
\end{abstract}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{fig/docsplit.pdf}
\caption{Overall framework of \our. The document is randomly divided into two exclusive subsets of sentences and the two subsets work as positive pairs for contrastive learning. Other instances in the same batch are used as negatives.
}
\label{overall}
\end{figure}

@@

To evaluate the quality of document embeddings, we conduct standard and few-shot text classification on five large document datasets involved in News and scientific articles. Moreover, to further validate the effectiveness of our method, we also conduct document retrieval on the AAN dataset which is designed for long document understanding evaluation.
The experimental results show that \our~with two kinds of model structures (i.e.,~BERT and Longformer) can both achieve significant improvements compared to state-of-the-art baselines.

@@

\subsection{DocSplit}
The critical problem in contrastive learning is how to construct positive pairs $(x, x^{+})$.
In representation learning for visual tasks~\cite{Chen2020ASF}, an effective solution is to take two random transformations of the same image (e.g.,~flipping, rotation).
Similarly, in language representations, previous works~\cite{Gao2021SimCSESC, Karpukhin2020DensePR, Meng2021COCOLMCA, Li2022UCTopicUC} apply augmentation techniques such as dropout, word deletion, reordering, and masking.

@@

\begin{table*}
    \centering
% \scalebox{0.7}{
% \setlength{\tabcolsep}{1.mm}{
    \input{fig/result.tex}
    % }}
    \caption{For all performance measures, larger numbers are better. Our pre-trained model achieves the best results in all cases. *INSTRUCTOR~\cite{Su2022OneEA} includes 20News in their pretraining datasets which causes data leakage in our experiments, hence INSTRUCTOR results on 20News are not comparable to others.}
    \vspace{-3mm}
    \label{results}
\end{table*}

@@

\subsection{Baselines}
We compare our pre-trained model to the baselines of two groups.
(1) Language models trained for general language understanding and sentence embeddings include BERT~\cite{Devlin2019BERTPO}, SimCSE~\cite{Gao2021SimCSESC}~\footnote{SimCSE can be viewed as an ablation model without document splitting.}, CT-BERT~\cite{Carlsson2021SemanticRW}, Contriever~\cite{Izacard2021UnsupervisedDI} and INSTRUCTOR~\cite{Su2022OneEA}. 
For a fair comparison, we also train a SimCSE with our pretraining dataset (SimCSE$_\mathrm{long}$).
(2) Transformers specified for long sequences include Longformer~\cite{Beltagy2020LongformerTL} and BigBird~\cite{Zaheer2020BigBT}.
We train two versions of \our~with BERT and Longformer (i.e.,~\our$_\mathrm{bert}$ and \our$_{\mathrm{long}}$) for comparison.
We do not include RoBERTa~\cite{Liu2019RoBERTaAR}, IS-BERT~\cite{Zhang2020AnUS} and GTR~\cite{Ni2021LargeDE} as our baselines because SimCSE and INSTRUCTOR achieve better results than these methods according to their papers.

@@

\textbf{Datasets.}
We use the following classic documents datasets to evaluate our method:
(1) Fake News Corpus~\footnote{https://github.com/several27/FakeNewsCorpus};
(2) 20NewsGroups~\cite{Lang1995NewsWeederLT};
(3) arXiv articles dataset~\footnote{https://www.kaggle.com/datasets/Cornell-University/arxiv};
(4) New York Times Annotated Corpus (NYT) ~\cite{sandhaus2008new}; and
(5) BBCNews~\footnote{http://mlg.ucd.ie/datasets/bbc.html}.
We do not use semantic textual similarity (STS) tasks~\cite{Agirre2012SemEval2012T6} because the sentences in these tasks are short which is not suitable to evaluate large document embeddings.

\textbf{Results.}
Table~\ref{results} shows the evaluation results on different datasets. 
Overall, we can see that \our~achieves the best performance over the 4 document datasets (except INSTRUCTOR on 20News due to its data leakage) and consistently improves the document embeddings with BERT and Longformer structures. 
Specifically, methods pretrained with contrastive objectives (i.e.,~CT-BERT, SimCSE, Contriever and INSTRUCTOR) outperform general language representations (i.e., BERT and LongFormer) which indicates contrastive objectives designed for text embeddings can largely improve the ability of language models to produce high-quality text embeddings. 
SimCSE pretrained with our document data (i.e.,~SimCSE$_{\mathrm{long}}$) has similar results as the original SimCSE which indicates simply increasing the length of pretraining text cannot improve document embeddings.
Under the same model structures, our model achieves $3.9\%$ and $9.4\%$ average macro-F1 improvements with BERT and Longformer structures respectively compared to SimCSE and Longformer. 
Compared to the state-of-the-art model INSTRUCTOR, our model (\our$_{\mathrm{bert}}$) can achieve $3.4\%$ improvements.
Hence, our contrastive learning method is effective for document embeddings.

@@

\textbf{Results.} 
Table~\ref{results} shows the results of few-shot text classification on these five datasets. 
We can see that, under the same model structure, \our~(i.e.,~\our$_{\mathrm{bert}}$ and \our$_{\mathrm{long}}$) achieves $12.0\%$ and $24.3\%$ macro-F1 improvements compared to SimCSE and Longformer respectively. 
Surprisingly, INSTRUCTOR achieves low performance on the NYT and BBCNews under few-shot settings.
%and achieves 0.028+0.082+2.03+0.769. 
These improvements are higher than standard text classification. 
Besides, we also compare the performance of different baselines (using BERT as a backbone) and \our$_{\mathrm{bert}}$ with different numbers of training instances on 20News. 
The results in Figure~\ref{few_shot} show the improvements from our method become larger as the number of training instances decreases indicating the importance of high-quality document embeddings for low-resource settings. 
Furthermore, our method achieves the best results under different numbers of training instances.

\subsection{Document Retrieval}
We conduct document retrieval to evaluate the ability to learn the similarity score between two vectors~\cite{Guo2016ADR} of documents. We follow the document retrieval experiment~\cite{Tay2020LongRA} and use the ACL Anthology Network (AAN)~\cite{Radev2009TheAA} dataset, which identifies if two papers have a citation link, a common setup used in long-form document matching~\cite{Jiang2019SemanticTM,Yang2020Beyond5T}.
Specifically, for all baselines, we use models to obtain document embeddings and finetune an MLP layer on two concatenated embeddings to predict if two documents have a citation link.

\textbf{Results.} 
\begin{table*}
    \centering
    \input{fig/document_rerival.tex}
    \caption{}
    \vspace{-3mm}
    \label{retrieval}
\end{table*}
Table~\ref{retrieval} shows the results of the document retrieval experiment on AAN dataset.
We can find that our method achieves the best results due to the effectiveness of our pretraining methods based on document splitting.

\section{Conclusion}
\label{sec:conclusion}
In this work, we propose an unsupervised contrastive learning framework for large document embeddings. 
Our paper provides a new method for large document data augmentation without any supervision and language models can get large-scale pretraining on any large documents. 
We conduct extensive experiments on text classification tasks under fully supervised and few-shot settings and a document retrieval task for further evaluation.
Results show that our pre-trained model greatly outperforms state-of-the-art text embeddings, especially when the training data is limited.

@@

\textbf{Text Embeddings}. Learning text embeddings is a fundamental task in natural language processing (NLP) which can be used in information retrieval~\cite{Thakur2021BEIRAH}, text similarity~\cite{Gao2021SimCSESC}, classification~\cite{Minaee2021DeepLB}, etc. Existing works focus on learning general text embeddings for different applications. SimCSE~\cite{Gao2021SimCSESC} augment training data by Dropout~\cite{Srivastava2014DropoutAS} to construct positive data for contrastive learning. Sentence-BERT~\cite{Reimers2019SentenceBERTSE} employs siamese and triplet network structures to compare the labeled sentences in the natural language inference (NLI) dataset. Contriever~\cite{Izacard2021UnsupervisedDI} learns text embeddings for information retrieval via contrastive learning considering multilingual corpus. INSTRUCTOR~\cite{Su2022OneEA} collects a massive dataset from 330 diverse NLP tasks and formulates all tasks into information retrieval tasks with prompts. 
Previous works either rely on massive manually annotated data~\cite{Su2022OneEA, Reimers2019SentenceBERTSE} or focus only on short text (i.e.,~sentences) embeddings~\cite{Gao2021SimCSESC}.
In this paper, we study unsupervised representation learning of document embeddings which usually have longer text than general sentence embeddings and this task brings new challenges for models to understand a document.

@@

% \subsection{Attention Weights}
% \label{app:attention}
% To explore the difference between \our~and other models, we analyze the attention weights of Transformers in different models on the NYT dataset (details in Appendix~\ref{app:attention}). The average weights of different kinds of words are shown in Figure~\ref{attentions}.  We can see that our model has more than $40\%$ higher attention weights on nouns compared to BERT and SimCSE.~\citet{martin-johnson-2015-efficient} shows nouns are more informative than other words in the document understanding. Hence, our pretraining method increases the attention weights of models on nouns which results in higher performance on long text classification.

% \begin{figure}
% \centering
% \includegraphics[width=0.8\linewidth]{fig/attentions.pdf}
% \caption{Attention weights from different models on the NYT dataset.}  
% \label{attentions}
% \end{figure}

% We compute the attention weights for Transformers as follows:
% (1) we first extract the attention weights between \texttt{[CLS]} token and all the other tokens;
% (2) we compute the averaged weights along different heads in multi-head attention;
% (3) the attention weights of the last layer in Transformers are used as the weights for words.
% Averaged values are computed for nouns, verbs, adjectives, and other words.
